---
title: "Assignment3"
author: "RF"
date: "3/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Multivariate linear models

In this assignment you will investigate the relation between different aspects of IQ and symptom severity in ASD (as measured by ADOS). The assignment is meant to make you practice on linear models, multiple predictors and interactions.

The data you will use is the outcome of the psychological testing of the children you analyzed in methods 3.
Data: https://www.dropbox.com/s/hom2qnmp9hyfwmu/Assignment3Data.csv?dl=0
The data consists of ChildID, gender, age, diagnosis (ASD), symptom severity (ADOS), several aspects of IQ (NonVerbal/Perceptual,Verbal,Social) and the visit at which the tests were administered. The questions should be answered by focusing on visit 1.

```{r}
p_load(tidyverse, rethinking)
setwd("~/Desktop/Computational Models/P3")

d1 = read.csv("Assignment3Data.csv")
#making a subset containing only values from visit 1 and children with ASD
d = subset(d1, Visit == 1 & ASD == 1)
```
The questions you will have to answer are:

1. Assess the relation between symptom severity and IQ (focus on visit 1 and children with ASD) and report the model, a plot of the model, a couple of lines describing the quality of the model and interpreting the results. P.S. Should you scale? P.P.S. Don't forget to motivate your priors. P.P.P.S. At least one plot for results and a plot for quality of each model (here and in the next questions) would be appreciated.
1.1. Verbal IQ and ADOS
```{r}
#creating model, using verbal IQ to predict ADOS
#|Diagnosis |  meanVIQ|    sdVIQ| meanNVIQ|   sdNVIQ|    ADOS| sdADOS|
#|:---------|--------:|--------:|--------:|--------:|-------:|------:|
#|ASD       | 17.31034| 7.450328| 26.89655| 5.677932| 14.2069|     NA|
#|TD        | 20.21875| 5.128883| 26.00000| 3.340852|  0.8750|     NA|

d$ScaleADOS = scale(d$ADOS)
d$ScaleVIQ = scale(d$VerbalIQ)
d$ScaleNVIQ = scale(d$NonVerbalIQ)
d$ScaleSIQ = scale(d$SocialIQ)

#from previous assignments we know that ADOS ranges from approx 6 - 21 in ASD children. 

#CHANGE THE PRIORS TO SCALED 
m1.1 <- map(
  alist(
    ScaleADOS ~ dnorm( mu , sigma ) , #assuming ADOS is normally distributed
    mu <- a + b*ScaleVIQ ,
    a ~ dnorm( 0 , 2 ) , #ADOS has a mean of approx 14 and ranges from 6-21 in the previous assignment
    b ~ dnorm( 0 , 2 ) , #we don't know much about the relationship, so putting a liberal prior
    sigma ~ dunif( 0 , 10 ) #setting a quite generous SD since we don't know much
  ) ,
  data=d )
precis(m1.1)
#precis(m1.1, corr = T) 

#plotting the raw data
ggplot(d, aes(ScaleVIQ, ScaleADOS)) +
  geom_point()

#extracting samples from the posterior
post.samples=extract.samples(m1.1)

#plotting with the uncertainty
#just to check if the model makes sense
plot(ScaleADOS ~ ScaleVIQ, d)
abline(a=coef(m1.1)['a'], b=coef(m1.1)['b'])
for ( i in 1:1000 )
abline( a=post.samples$a[i] , b=post.samples$b[i] , col=col.alpha(rangi2, 0.5) )

# define sequence of weights to compute predictions for 
# these values will be on the horizontal axis
ScaleVIQ.seq <- seq( from=-2 , to=2.5 , by=0.1 )

# use link to compute mu for each sample from posterior and for each value in scaleVIQ.seq
mu <- link( m1.1 , data=data.frame(ScaleVIQ=ScaleVIQ.seq) )


# use type="n" to hide raw data
plot( ScaleADOS ~ ScaleVIQ , d , type="n" )
# loop over samples and plot each mu value
for ( i in 1:nrow(mu) )
points( ScaleVIQ.seq , mu[i,] , pch=16 , col=col.alpha(rangi2,0.1) )

#summarize the distribution of mu
mu.mean <- apply( mu , 2 , mean )
mu.HPDI <- apply( mu , 2 , HPDI , prob=0.89 )

# plot raw data
# fading out points to make line and interval more visible
plot( ScaleADOS ~ ScaleVIQ , data=d , col=col.alpha(rangi2,0.5) )
# plot the MAP line, aka the mean mu for each weight
lines( ScaleVIQ.seq , mu.mean )
# plot a shaded region for 89% HPDI
shade( mu.HPDI , ScaleVIQ.seq )

#simulating ADOS values for the specific VerbalIQ values in the previously defined sequence
sim.ados <- sim( m1.1 , data=list(ScaleVIQ=ScaleVIQ.seq) )
#summarizing the distribution of the simulation
ados.HPDI <- apply( sim.ados , 2 , HPDI , prob=0.89 )

# plot raw data
plot( ScaleADOS ~ ScaleVIQ , d , col=col.alpha(rangi2,0.5) )
# draw MAP line
lines( ScaleVIQ.seq , mu.mean )
# draw HPDI region for line
shade( mu.HPDI , ScaleVIQ.seq )
# draw PI region for simulated heights
shade( ados.HPDI , ScaleVIQ.seq )

```


#Function to create the plot
```{r}
#Plot to make the plot with MAP line, HPDI for the line and HPDI for the model with all the uncertainty
#takes a model, the x value as a string (ie. in ''), y value in string and the dataframe.
fancypantsplot <- function(model, x, y, dataframe) {
 #extracting data from the dataframe 
  xd <- dplyr::select(dataframe, x) 
  yd <- dplyr::select(dataframe, y)
  
  #calculate the prior data - the next 4 lines basically recreate the link function (see page 107)
  post <- extract.samples(model) 
  mu.link <- function(x) post$a + post$b*x 
  x.seq <- seq( from=min(xd) , to=max(xd) , by=0.1 ) 
  mu <- sapply( x.seq , mu.link ) 
  #summarize the distribution of mu
  mu.mean <- apply( mu , 2 , mean ) 
  mu.HPDI <- apply( mu , 2 , HPDI , prob=0.89 ) 
  
  temp_list = list(x.seq); names(temp_list) <- x  #naming the vector in the list (since the sim function apparently  requires it - thank Kenneth)
  sim.x <- sim( model , data=temp_list ) 
  x.HPDI <- apply( sim.x , 2 , HPDI , prob=0.89 )
  
  #creating dataframe containing only the x and y values
  new_df = data.frame(xd, yd)
  #giving the columns new names that can be called in the plot function
  names(new_df) <- c('x1', 'y1')
  # plot raw data
  plot( y1 ~ x1 , new_df, col=col.alpha(rangi2,0.5) )
  # draw MAP line
  lines( x.seq , mu.mean )
  # draw HPDI region for line
  shade( mu.HPDI , x.seq )
  # draw HPDI region for simulated heights
  shade( x.HPDI , x.seq )
}

#IN ACTION!
fancypantsplot(m1.1, 'ScaleVIQ', 'ScaleADOS', d)

```


#Kenneths psykopat function til at lave plottet (virker ikke for mig)
```{r}

gggplot = function(model, x = NULL , y = NULL, dataframe){ 
   #where: 
  #model is a model resulting from MAP 
  #x is the x variable as a string (can be left unspecified) 
  #y is the y variable as a string (can be left unspecified) 
  #dataframe is the dataframe from which the model is trained upon 
  #requires packages: dplyr, ggplot2, stringr and rethinking
  
  if (is.null(y) == T){ 
    #if y isn't specified - extract y 
    temp <- flist_untag(model@formula)[[1]] 
    y <- as.character(temp[[2]])
    } 
  if (is.null(x) == T){ 
    #if x isn't specified - extract x 
      temp <- flist_untag(model@formula)[[2]] 
      x <- gsub(" ","",gsub(".*\\*","",as.character(temp[[3]])[3])) 
  }
  
   #extracting data from the dataframe 
  xd <- dplyr::select(dataframe, x) 
  yd <- dplyr::select(dataframe, y)
  
   #calculate the prior data 
  post <- extract.samples(model) 
  #posterior from samples 
  mu.link <- function(x) post$a + post$b*x 
  VIQ.seq <- seq( from=min(xd) , to=max(xd) , by=0.1 ) 
  mu <- sapply( VIQ.seq , mu.link ) 
  mu.mean <- apply( mu , 2 , mean ) 
  mu.HPDI <- apply( mu , 2 , HPDI , prob=0.89 ) 
  temp_list = list(VIQ.seq); names(temp_list) <- x 
  #naming the vector in the list (since the sim function requires it) 
  sim.VIQ <- sim( model , data=temp_list ) 
  VIQ.PI <- apply( sim.VIQ , 2 , PI , prob=0.89 )

  #transform the data from ggplot 
    #for making the line 
  temp_df <- data.frame(VIQ.seq = VIQ.seq, mu.mean = mu.mean) 
  #for making the HDPI 
  temp_matrix <- t(data.frame(mu.HPDI = mu.HPDI)) 
  #t() is a tranpose function which returns a matrix 
  temp_df1 = data.frame(lower = temp_matrix[,1], upper = temp_matrix[,2], mu.mean = mu.mean, VIQ.seq = VIQ.seq)
   #for making the PI 
  temp_matrix <- t(data.frame(VIQ.PI = VIQ.PI)) 
  temp_df2 = data.frame(lower = temp_matrix[,1], upper = temp_matrix[,2], mu.mean = mu.mean, VIQ.seq = VIQ.seq) 
  #for the points 
  dataframe = data.frame(xd, yd) 
  #taking the data to new column as to chose the name (maintaining old name for the names on the plot) 
  dataframe$c1 = dataframe[,1] 
  dataframe$c2 = dataframe[,2]
  
   #make the plot 
  grob = ggplot(dataframe, aes(x = c1, y = c2)) + 
    geom_ribbon(data = temp_df2, aes(x = VIQ.seq, y = mu.mean, ymin = lower, ymax = upper), fill = "grey70", alpha = 2/3) + 
    geom_ribbon(data = temp_df1, aes(x = VIQ.seq, y = mu.mean, ymin = lower, ymax = upper), fill = "grey83", alpha = 2/3) + 
    geom_line(data = temp_df, aes(x = VIQ.seq, y = mu.mean))+ geom_point(color = "navyblue", shape = 1, size = 1.5, alpha = 2/3)+ 
    xlab(colnames(dataframe)[1]) + 
    ylab(colnames(dataframe)[2]) 
  #using the names for the axis return(grob) 
  } #go go gadget fancy plot

gggplot(m1.1, dataframe = d)

```


1.2. Non Verbal IQ and ADOS
1.3. Social IQ and ADOS

2. Do the different aspects of IQ account for different portions of the variance in ADOS? 
2.1. Does it make sense to have all IQ measures in the same model? First write a few lines answering the question and motivating your answer, including a discussion as to what happens when you put all of them in the same model. Then build a model following your answer. If your answer is "no", you are not free, you still have to answer: are there alternative ways of answering the question?
2.2. Build the model, assess its quality, write a few lines interpreting the results.

3. Let's now include also the TD children. Does it make sense to ask whether IQ and ADOS are related? Motivate your answer. In any case, if you wanted to build a model to answer that question, which model would you build? Run the model, assess its quality, write a few lines interpreting the results.

4. Let's discuss contents:
4.1. You have three scores for IQ, do they show shared variance? Is that the same in TD and ASD? What does that tell us about IQ?
4.2. You have explored the relation between IQ and ADOS. How do you explain that relation from a cognitive perspective? N.B. You can present alternative hypotheses.

5. Bonus questions: Including measurement errors. 
5.1. Let's assume that ADOS has a measurement error of 1. How would you model that in any of the previous models? 
5.2. We know that IQ has an estimated measurement error of 2.12. How would you include that? 


